---
title: "Sams_paper_clustering"
author: "Devika Agarwal"
date: 'Last update: `r date()`'
output:
  html_document:
    theme: cerulean
    number_sections: true
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: yes
      toc_depth: 6
---
<style type="text/css">

body{ /* Normal  */
      font-size: 12px;
  }
td {  /* Table  */
  font-size: 10px;
}
h1.title {
  font-size: 18px;
  color: DarkRed;
}
h1 { /* Header 1 */
  font-size: 16px;
  color: DarkBlue;
}
h2 { /* Header 2 */
    font-size: 16px;
  color: DarkBlue;
}
h3 { /* Header 3 */
  font-size: 14px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```



```{r pacakges}
library(Seurat)
library(plyr)
library(tidyverse)
library(ggplot2)
library(psych)
library(tidyr)
library(randomcoloR)
```



# Clustering of Organoid + Human Cerebellar data

```{r}
Aldinger_organoid <- readRDS(file="./20200918KALoadganoidsSCTBest.rds")
Idents(object = Aldinger_organoid) <- "old.ident"

Aldinger_organoid$age <- Aldinger_organoid$age %>% replace_na("D90")
Aldinger_organoid$age <- as.factor(Aldinger_organoid$age)
Aldinger_organoid$celltype_age <- paste(Aldinger_organoid$age,"_",Aldinger_organoid$old.ident,sep="")

Avg_Aldinger_organoid <-  AverageExpression(Aldinger_organoid, return.seurat = T,)

Avg_log_int_KA_Nay <- GetAssayData(Avg_Aldinger_organoid ,slot = "data", assay = "integrated")

Avg_scaled_int_KA_Nay <- GetAssayData(Avg_Aldinger_organoid, slot = "scale.data", assay = "integrated")

Avg_raw_int_KA_Nay <- GetAssayData(Avg_Aldinger_organoid, slot = "counts", assay = "integrated")
```

# heatmap/Clustering of integrated cell type data (Oragnoid + KA human cerebellar data)

## Average scaled integrated seurat data

```{r, fig.height=7, fig.width=8}
library(pheatmap)
pearson_corr <- cor(Avg_scaled_int_KA_Nay, method = "pearson")
celltype_scale_KA <- pheatmap::pheatmap(mat= pearson_corr, clustering_method = "ward.D2", fontsize = 5)
```

## ComBat Correction of Average logged data (non parametric combat corrected)

```{r}
metadata <- Avg_Aldinger_organoid@meta.data 
metadata$study <- c( rep("KA", times=21), rep("Nayler",times=12))
metadata$study <- as.factor(metadata$study)
library(sva)
combat_avg_log_KA <- ComBat(dat = as.matrix(Avg_log_int_KA_Nay), batch = metadata$study, par.prior = F)
```



```{r, fig.height=7, fig.width=8}
pearson_corr <- cor(combat_avg_log_KA, method = "pearson")
ann_col <- data.frame(rownames(metadata),metadata$study)
colnames(ann_col)[2] <- "Study"
colnames(ann_col)[1] <- "row"
ann_col<- column_to_rownames(ann_col, var = "row")
celltype_combat_KA_1 <- pheatmap::pheatmap(mat= pearson_corr, clustering_method = "ward.D2", annotation_col = ann_col, fontsize = 6)
```


## ComBat Correction of Average logged data (parametric combat corrected)

```{r,fig.height=7, fig.width=8}
combat_avg_log_KA <- ComBat(dat = as.matrix(Avg_log_int_KA_Nay), batch = metadata$study, par.prior = T )
pearson_corr <- cor(combat_avg_log_KA, method = "pearson")
ann_col <- data.frame(rownames(metadata),metadata$study)
colnames(ann_col)[2] <- "Study"
colnames(ann_col)[1] <- "row"
ann_col<- column_to_rownames(ann_col, var = "row")
celltype_combat_KA_2 <- pheatmap::pheatmap(mat= pearson_corr, clustering_method = "ward.D2", annotation_col = ann_col, fontsize = 5)
```


## PCA of Average integrated data (log or scaled)

```{r, fig.height=7,fig.width=10}
metadata$celltype <- rownames(metadata)
pc_KA_avg <- prcomp(t(combat_avg_log_KA), scale. = F)

metadata$pc1 <- pc_KA_avg$x[, 1]
metadata$pc2 <-  pc_KA_avg$x[, 2]
metadata$pc3 <- pc_KA_avg$x[, 3]
metadata$pc4 <- pc_KA_avg$x[, 4]
metadata$pc5 <- pc_KA_avg$x[, 5]
metadata$pc6 <- pc_KA_avg$x[, 6]
  pve_pc_KAavg <- summary(pc_KA_avg)$importance[2, ]
  
  #percent_var <- round( signif((pve[2]) * 100, 4), 2)
#library(RColorBrewer)
#colourCount = length(unique(metadata$celltype))
#colourCount2= length(unique(plotdata$Batch))
#getPalette = colorRampPalette(brewer.pal(11, "Set1"))
  
colour <- randomColor(count = 33)
  
library(ggrepel)


p <- ggplot(data = metadata, aes(x=pc1, y=pc2, color= celltype, shape=study, label=celltype))
p <- p + geom_point(alpha = 0.6, size=8)
p <- p + scale_colour_manual(values = colour)
p <- p + xlab(paste0("PC1 (", signif((pve_pc_KAavg[1]) * 100, 4), "%)"))
p <- p + ylab(paste0("PC2 (",signif((pve_pc_KAavg[2]) * 100, 4), "%)"))
p <- p + geom_text_repel(point.padding = NA, size=4)
p <- p + geom_hline(yintercept=0, linetype="dashed", color = "gray")
p <- p + geom_vline(xintercept=0, linetype="dashed", color = "gray")
p <- p + guides(colour = guide_legend(override.aes = list(alpha=1))) + theme_bw()
#p <- p + theme(legend.position = "none") 
p


p <- ggplot(data = metadata, aes(x=pc2, y=pc3, color= celltype, shape=study,label=celltype))
p <- p + geom_point(alpha = 0.6, size=8)
#p <- p + scale_colour_manual(values = getPalette(colourCount))
p <- p + xlab(paste0("PC2 (", signif((pve_pc_KAavg[2]) * 100, 4), "%)"))
p <- p + ylab(paste0("PC3 (",signif((pve_pc_KAavg[3]) * 100, 4), "%)"))
p <- p + geom_text_repel(point.padding = NA, size=4)
p <- p + geom_hline(yintercept=0, linetype="dashed", color = "gray")
p <- p + geom_vline(xintercept=0, linetype="dashed", color = "gray")
p <- p + guides(colour = guide_legend(override.aes = list(alpha=1))) + theme_bw()
#p <- p + theme(legend.position = "none") 
p

p <- ggplot(data = metadata, aes(x=pc3, y=pc4, color= celltype, shape=study,label=celltype))
p <- p + geom_point(alpha = 0.6, size=8)
#p <- p + scale_colour_manual(values = getPalette(colourCount))
p <- p + xlab(paste0("PC3 (", signif((pve_pc_KAavg[3]) * 100, 4), "%)"))
p <- p + ylab(paste0("PC4 (",signif((pve_pc_KAavg[4]) * 100, 4), "%)"))
p <- p + geom_text_repel(point.padding = NA, size=4)
p <- p + geom_hline(yintercept=0, linetype="dashed", color = "gray")
p <- p + geom_vline(xintercept=0, linetype="dashed", color = "gray")
p <- p + guides(colour = guide_legend(override.aes = list(alpha=1))) + theme_bw()
#p <- p + theme(legend.position = "none") 
p


```


# Aldinger + Organoid (celltype + age)

```{r}
Idents(object = Aldinger_organoid) <- "celltype_age"
Avg_Aldinger_organoid <-  AverageExpression(Aldinger_organoid, return.seurat = T,)

Avg_log_int_KA_Nay <- GetAssayData(Avg_Aldinger_organoid ,slot = "data", assay = "integrated")

Avg_scaled_int_KA_Nay <- GetAssayData(Avg_Aldinger_organoid, slot = "scale.data", assay = "integrated")
```


## Average scaled integrated seurat data

```{r, fig.height=13, fig.width=13}
library(pheatmap)
pearson_corr <- cor(Avg_scaled_int_KA_Nay, method = "pearson")
celltype_scale_KA <- pheatmap::pheatmap(mat= pearson_corr, clustering_method = "ward.D2", fontsize = 5,filename = "./Aldinger_Nayler_age_celltype_integrated_scaled.pdf",height = 16, width = 16,border_color = NA, fontsize_row = 5,fontsize_col = 5)
celltype_scale_KA
```

## ComBat Correction of Average logged data (non parametric combat corrected)

```{r}
metadata <- Avg_Aldinger_organoid@meta.data 
metadata$study <- c( rep("KA", times=196), rep("Nayler",times=12))
metadata$study <- as.factor(metadata$study)
library(sva)
combat_avg_log_KA <- ComBat(dat = as.matrix(Avg_log_int_KA_Nay), batch = metadata$study, par.prior = F)
```



```{r, fig.height=13, fig.width=13}
pearson_corr <- cor(combat_avg_log_KA, method = "pearson")
ann_col <- data.frame(rownames(metadata),metadata$study)
colnames(ann_col)[2] <- "Study"
colnames(ann_col)[1] <- "row"
ann_col<- column_to_rownames(ann_col, var = "row")
celltype_combat_KA_1 <- pheatmap::pheatmap(mat= pearson_corr, clustering_method = "ward.D2", annotation_col = ann_col, fontsize = 5,filename = "./Aldinger_Nayler_age_celltype_combat_non_parametric.pdf",height = 16, width = 16,border_color = NA, fontsize_row = 5,fontsize_col = 5)
```

## ComBat Correction of Average logged data (parametric combat corrected)

```{r,fig.height=13, fig.width=13}
combat_avg_log_KA <- ComBat(dat = as.matrix(Avg_log_int_KA_Nay), batch = metadata$study, par.prior = T )
pearson_corr <- cor(combat_avg_log_KA, method = "pearson")
ann_col <- data.frame(rownames(metadata),metadata$study)
colnames(ann_col)[2] <- "Study"
colnames(ann_col)[1] <- "row"
ann_col<- column_to_rownames(ann_col, var = "row")
celltype_combat_KA_2 <- pheatmap::pheatmap(mat= pearson_corr, clustering_method = "ward.D2", annotation_col = ann_col, fontsize = 5,border_color = NA, fontsize_row = 5,fontsize_col = 5)
```

## PCA of Average integrated data (log or scaled)

```{r, fig.height=7,fig.width=10}
metadata$celltype <- rownames(metadata)
pc_KA_avg <- prcomp(t(combat_avg_log_KA), scale. = F)

metadata$pc1 <- pc_KA_avg$x[, 1]
metadata$pc2 <-  pc_KA_avg$x[, 2]
metadata$pc3 <- pc_KA_avg$x[, 3]
metadata$pc4 <- pc_KA_avg$x[, 4]
metadata$pc5 <- pc_KA_avg$x[, 5]
metadata$pc6 <- pc_KA_avg$x[, 6]
  pve_pc_KAavg <- summary(pc_KA_avg)$importance[2, ]
  
  #percent_var <- round( signif((pve[2]) * 100, 4), 2)
#library(RColorBrewer)
#colourCount = length(unique(metadata$celltype))
#colourCount2= length(unique(plotdata$Batch))
#getPalette = colorRampPalette(brewer.pal(11, "Set1"))
  
colour <- randomColor(count = 33)
  
library(ggrepel)


p <- ggplot(data = metadata, aes(x=pc1, y=pc2, color= celltype, shape=study, label=celltype))
p <- p + geom_point(alpha = 0.6, size=8)
#p <- p + scale_colour_manual(values = colour)
p <- p + xlab(paste0("PC1 (", signif((pve_pc_KAavg[1]) * 100, 4), "%)"))
p <- p + ylab(paste0("PC2 (",signif((pve_pc_KAavg[2]) * 100, 4), "%)"))
p <- p + geom_text_repel(point.padding = NA, size=4)
p <- p + geom_hline(yintercept=0, linetype="dashed", color = "gray")
p <- p + geom_vline(xintercept=0, linetype="dashed", color = "gray")
p <- p + guides(colour = guide_legend(override.aes = list(alpha=1))) + theme_bw()
p <- p + theme(legend.position = "none") 
p


p <- ggplot(data = metadata, aes(x=pc2, y=pc3, color= celltype, shape=study,label=celltype))
p <- p + geom_point(alpha = 0.6, size=8)
#p <- p + scale_colour_manual(values = getPalette(colourCount))
p <- p + xlab(paste0("PC2 (", signif((pve_pc_KAavg[2]) * 100, 4), "%)"))
p <- p + ylab(paste0("PC3 (",signif((pve_pc_KAavg[3]) * 100, 4), "%)"))
p <- p + geom_text_repel(point.padding = NA, size=4)
p <- p + geom_hline(yintercept=0, linetype="dashed", color = "gray")
p <- p + geom_vline(xintercept=0, linetype="dashed", color = "gray")
p <- p + guides(colour = guide_legend(override.aes = list(alpha=1))) + theme_bw()
p <- p + theme(legend.position = "none") 
p

p <- ggplot(data = metadata, aes(x=pc3, y=pc4, color= celltype, shape=study,label=orig.ident))
p <- p + geom_point(alpha = 0.6, size=8)
#p <- p + scale_colour_manual(values = getPalette(colourCount))
p <- p + xlab(paste0("PC3 (", signif((pve_pc_KAavg[3]) * 100, 4), "%)"))
p <- p + ylab(paste0("PC4 (",signif((pve_pc_KAavg[4]) * 100, 4), "%)"))
p <- p + geom_text_repel(point.padding = NA, size=4)
p <- p + geom_hline(yintercept=0, linetype="dashed", color = "gray")
p <- p + geom_vline(xintercept=0, linetype="dashed", color = "gray")
p <- p + guides(colour = guide_legend(override.aes = list(alpha=1))) + theme_bw()
p <- p + theme(legend.position = "none") 
p

write.table(metadata, "KA_Nayler_celltype_age_avg_metadata_pc_combat_corrected.txt", sep="\t", quote=F)

```

# Plot only PCs from KA and all of NAyler celltypes on PCA

```{r}
metadata_2 <- read.table(file="./KA_Nayler_PC_celltype_age_avg_metadata_pc_combat_corrected.txt", sep="\t", header=T, row.names = 1)
p <- ggplot(data = metadata_2, aes(x=pc1, y=pc2, color= celltype, shape=study, label=celltype))
p <- p + geom_point(alpha = 0.6, size=8)
#p <- p + scale_colour_manual(values = colour)
p <- p + xlab(paste0("PC1 (", signif((pve_pc_KAavg[1]) * 100, 4), "%)"))
p <- p + ylab(paste0("PC2 (",signif((pve_pc_KAavg[2]) * 100, 4), "%)"))
p <- p + geom_text_repel(point.padding = NA, size=4)
p <- p + geom_hline(yintercept=0, linetype="dashed", color = "gray")
p <- p + geom_vline(xintercept=0, linetype="dashed", color = "gray")
p <- p + guides(colour = guide_legend(override.aes = list(alpha=1))) + theme_bw()
p <- p + theme(legend.position = "none") 
p


p <- ggplot(data = metadata_2, aes(x=pc2, y=pc3, color= celltype, shape=study,label=celltype))
p <- p + geom_point(alpha = 0.6, size=8)
#p <- p + scale_colour_manual(values = getPalette(colourCount))
p <- p + xlab(paste0("PC2 (", signif((pve_pc_KAavg[2]) * 100, 4), "%)"))
p <- p + ylab(paste0("PC3 (",signif((pve_pc_KAavg[3]) * 100, 4), "%)"))
p <- p + geom_text_repel(point.padding = NA, size=4)
p <- p + geom_hline(yintercept=0, linetype="dashed", color = "gray")
p <- p + geom_vline(xintercept=0, linetype="dashed", color = "gray")
p <- p + guides(colour = guide_legend(override.aes = list(alpha=1))) + theme_bw()
p <- p + theme(legend.position = "none") 
p

p <- ggplot(data = metadata_2, aes(x=pc3, y=pc4, color= celltype, shape=study,label=celltype))
p <- p + geom_point(alpha = 0.6, size=8)
#p <- p + scale_colour_manual(values = getPalette(colourCount))
p <- p + xlab(paste0("PC3 (", signif((pve_pc_KAavg[3]) * 100, 4), "%)"))
p <- p + ylab(paste0("PC4 (",signif((pve_pc_KAavg[4]) * 100, 4), "%)"))
p <- p + geom_text_repel(point.padding = NA, size=4)
p <- p + geom_hline(yintercept=0, linetype="dashed", color = "gray")
p <- p + geom_vline(xintercept=0, linetype="dashed", color = "gray")
p <- p + guides(colour = guide_legend(override.aes = list(alpha=1))) + theme_bw()
p <- p + theme(legend.position = "none") 
p
```


```{r}
sessionInfo()
```

