---
title: "Sams_paper_clustering_2"
author: "Devika Agarwal"
date: 'Last update: `r date()`'
output:
  html_document:
    theme: cerulean
    number_sections: true
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: yes
      toc_depth: 6
---
<style type="text/css">

body{ /* Normal  */
      font-size: 12px;
  }
td {  /* Table  */
  font-size: 10px;
}
h1.title {
  font-size: 18px;
  color: DarkRed;
}
h1 { /* Header 1 */
  font-size: 16px;
  color: DarkBlue;
}
h2 { /* Header 2 */
    font-size: 16px;
  color: DarkBlue;
}
h3 { /* Header 3 */
  font-size: 14px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```



```{r pacakges}
library(Seurat)
library(plyr)
library(tidyverse)
library(ggplot2)
library(psych)
library(tidyr)
library(randomcoloR)
```

# Load in data

```{r load_data}
Human_mouse <- readRDS(file = "./20200924CarterRNAslotSCTmouseganoidsFIG.rds")

Human_mouse$celltype_int <-  paste(Human_mouse$old.ident,Human_mouse$cell_type,sep="")

Human_mouse$celltype_int <- gsub("NA","", Human_mouse$celltype_int)
#Human_mouse$sample_date <- as.factor(Human_mouse$sample_date)
Human_mouse$sample_date <- Human_mouse$sample_date %>% replace_na("D90")
Human_mouse$celltype_time <- paste(Human_mouse$sample_date,"_",Human_mouse$celltype_int,sep="")
Human_mouse$celltype_int <- paste(Human_mouse$celltype_int,"_",Human_mouse$dataset,sep="")
Idents(Human_mouse) <- "celltype_int"

metadata <- Human_mouse@meta.data
Avg_human_mouse <-  AverageExpression(Human_mouse, return.seurat = T,)

Avg_log_int <- GetAssayData(Avg_human_mouse, slot = "data", assay = "integrated")

Avg_scaled_int <- GetAssayData(Avg_human_mouse, slot = "scale.data", assay = "integrated")
```


# heatmap/Clustering of integrated cell type data (Oragnoid_Carter data)

```{r, fig.height=7, fig.width=8}
library(pheatmap)
pearson_corr <- cor(Avg_scaled_int, method = "pearson")
celltype_scale <- pheatmap::pheatmap(mat= pearson_corr, clustering_method = "ward.D2", fontsize = 5)
```

## ComBat Correction of Average logged data (non parametric combat corrected)

```{r}
metadata <- Avg_human_mouse@meta.data 
metadata$study <- c( rep("Nayler", times=12), rep("Carter",times=15))
metadata$study <- as.factor(metadata$study)
library(sva)
combat_avg_log_Carter <- ComBat(dat = as.matrix(Avg_log_int), batch = metadata$study, par.prior = F)
```

```{r, fig.height=7, fig.width=8}
pearson_corr <- cor(combat_avg_log_Carter, method = "pearson")
ann_col <- data.frame(rownames(metadata),metadata$study)
colnames(ann_col)[2] <- "Study"
colnames(ann_col)[1] <- "row"
ann_col<- column_to_rownames(ann_col, var = "row")
celltype_combat_KA_1 <- pheatmap::pheatmap(mat= pearson_corr, clustering_method = "ward.D2", annotation_col = ann_col, fontsize = 6)
```


## ComBat Correction of Average logged data (parametric combat corrected)

```{r}
metadata <- Avg_human_mouse@meta.data 
metadata$study <- c( rep("Nayler", times=12), rep("Carter",times=15))
metadata$study <- as.factor(metadata$study)
library(sva)
combat_avg_log_Carter <- ComBat(dat = as.matrix(Avg_log_int), batch = metadata$study, par.prior = T)
```

```{r, fig.height=7, fig.width=8}
pearson_corr <- cor(combat_avg_log_Carter, method = "pearson")
ann_col <- data.frame(rownames(metadata),metadata$study)
colnames(ann_col)[2] <- "Study"
colnames(ann_col)[1] <- "row"
ann_col<- column_to_rownames(ann_col, var = "row")
celltype_combat_KA_1 <- pheatmap::pheatmap(mat= pearson_corr, clustering_method = "ward.D2", annotation_col = ann_col, fontsize = 6)
```









# Developmental time + cell type clustering of Organoid : Carter data

```{r}
Idents(object = Human_mouse) <- "celltype_time"
Avg_human_mouse <-  AverageExpression(Human_mouse, return.seurat = T,)

Avg_log_int <- GetAssayData(Avg_human_mouse, slot = "data", assay = "integrated")

Avg_scaled_int <- GetAssayData(Avg_human_mouse, slot = "scale.data", assay = "integrated")
```

```{r,fig.height=13, fig.width=13}
pearson_corr <- cor(Avg_scaled_int, method = "pearson")
celltype_devtime <- pheatmap::pheatmap(mat= pearson_corr, clustering_method = "ward.D2", fontsize = 6, treeheight_col = 30, treeheight_row = 30)
```



```{r}
sessionInfo()
```

